---
title: "DATA607_Project2"
author: "Pei-Ming Chen"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
---

```{r load-packages, message=FALSE}
library(tidyverse)
library(openintro)
library(dplyr)
library(ggplot2)
```

### Dataset- Income 

The database contains 32,000 records on US Household Income Statistics & Geo Locations. The field description of the database is available on Kaggle.com. Income is a vital element when determining both quality and socioeconomic features of a given geographic location. The uploaded data was derived from over +36,000 files and covers 348,893 location records.

Importing the Data

```{r }
HI <- "https://raw.githubusercontent.com/PMCformosa/DATA607_Project-2/main/kaggle_income.csv"

Household_Income <- read.csv(file = HI, header =TRUE, sep=",")
head(Household_Income)
```


About the Dataset
The dataset was taken from Kaggle.com which was originally taken from an article called "US Household Income Statistics" Datasets
https://www.kaggle.com/datasets/goldenoakresearch/us-household-income-stats-geo-locations

```{r}
glimpse(Household_Income)

```

Data Wrangling
```{r}
colnames(Household_Income)
```

```{r}
colnames(Household_Income) <- c("Location_ID", "State_Code","State_Name","Abb_State_name","County", "City ", "Geo_location","Type","Primary","Zip_Code", "Area_Code","Area_Square","Water_area","Mean_H_Income","Sd_H_Income","Mean", "Median", "Stdev" , "sum_w" )
colnames(Household_Income) 

```

```{r}
glimpse(Household_Income)
```

```{r}
summary(Household_Income)

DF_H_Income <- Household_Income %>% as_tibble()

DF_H_Income

```

### Reshaping dataset

 Columns of County, Geo_location, and Type are collapsed into one column named Location and nother one with values on called District_type. Also , Area_Square, Water_area, Zip_Code are collapsed into the columns called Area and Area_size. The whole data frame looks much cleaner and consolidated.  

```{r}
DF_H_Income2 <- DF_H_Income %>%  pivot_longer(
    cols = c(County,Geo_location, Type, Primary), 
    names_to = "Location", 
    values_to = "District_type")

```


```{r}
DF_H_Income3 <- DF_H_Income2 %>%  pivot_longer(
    cols = c(Area_Square, Water_area, Zip_Code), 
    names_to = "Area", 
    values_to = "Area_size")

DF_H_Income3 
```

Counting possible missing Values 
```{r}
sum(is.na(DF_H_Income3)) 
```

### Plot the graph of average income of each state

Significantl , the highest average income is observed in California state 
```{r}
ggplot(DF_H_Income3, aes(x=Mean, y=Abb_State_name)) + geom_col()
```
The relationship of size of Area and its income level is shown in the graph below. 
```{r}
Household_Income %>% 
ggplot(aes(x = Area_Square, y = Mean)) +
  geom_point(color= "purple")
```

Puerto rico was chose to study here for its relatively low mean value in income . Its area-sdize and income relashionship is shown below. 
```{r}
PR_data <- DF_H_Income3[DF_H_Income3$State_Name == 'Puerto Rico' ,]
PR_data

summary(PR_data)

PR_dataB <- PR_data[order(PR_data$Mean, decreasing = TRUE),]

PR3 <- select(PR_dataB, "Mean")
PR3

ggplot(PR_data, aes(x = Area_size, y= Mean)) + geom_point(color="blue")

```


```{r}
CA_data <- DF_H_Income3[DF_H_Income3$State_Name == 'California' ,]
CA_data 

summary(CA_data)

CA_data2 <- CA_data[order(CA_data$Mean, decreasing = TRUE),]
CA_data2

CA3 <- select(CA_data2, "Mean")
CA3

ggplot(CA_data, aes(x = Area_size, y= Mean )) + geom_point(color="green")
```

